---
title: "DADA2_Assignment3"
author: "Claudia"
date: "08/11/2021"
output: html_document
---

#Set working directory to where your files are saved/where you want to work. 
```{r}
setwd ("~/Desktop/ENVS_BINF/Assignment 3/")
```

#Load in libraries required for DADA2 
#DADA2 must already be installed
```{r}
library(dada2); packageVersion("dada2")
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
library(ggplot2); packageVersion("ggplot2")
```

#Change path to where your data is located. 
```{r}
path <- "~/Desktop/ENVS_BINF/Assignment 3/Sequences" 
list.files(path)
```

#Tell DADA what the forward and reverse reads are. You need to know what the file format names are for your file (change the pattern="" section). 
```{r}
# Forward and reverse fastq filenames have format: SAMPLENAME_R1_001.fastq and SAMPLENAME_R2_001.fastq
fnFs <- sort(list.files(path, pattern="_R1_001.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_001.fastq", full.names = TRUE))
# Extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
sample.names
```

#visualize the quality profiles. Manually stop and look at the data. 
```{r}
plotQualityProfile(fnFs[1:2])
plotQualityProfile(fnRs[1:2])
```

#Assign filenames to the filtered files 
```{r}
# Place filtered files in filtered/ subdirectory
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
names(filtFs) <- sample.names
names(filtRs) <- sample.names
```

#here we actually filter and we choose the length at which to truncate the reads. truncLen=c(forward read cutoff#, reverse read cutoff#)
```{r}
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(260,210),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=TRUE) # On Windows set multithread=FALSE
head(out)
```

#learn the error rates of your reads. Visualize the estimated error rates.
```{r}
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
```

#how many different sequences do we have in each Sample?
```{r}
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)
dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
```

```{r}
dadaFs
```

#merge paired-end reads 
```{r}
mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
mergers
```

## Inspect distribution of sequence lengths
```{r}
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
table(nchar(getSequences(seqtab)))
```

#Sequences that are much longer or shorter than expected may be the result of non-specific priming. You can remove non-target-length sequences from your sequence table 
```{r}
seqtab2 <- seqtab[,nchar(colnames(seqtab)) %in% 396:420]
```

#Remove chimeras
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab2, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
sum(seqtab.nochim)/sum(seqtab)
```

#As a final check of our progress, we’ll look at the number of reads that made it through each step in the pipeline
```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
# If processing a single sample, remove the sapply calls: e.g. replace sapply(dadaFs, getN) with getN(dadaFs)
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
```

#Assign taxonomy using the Silva Database 
```{r}
taxa <- assignTaxonomy(seqtab.nochim, "~/Desktop/ENVS_BINF/DADA2/silva_nr99_v138.1_train_set.fa", multithread=TRUE)
```

```{r}
taxa.print <- taxa # Removing sequence rownames for display only
rownames(taxa.print) <- NULL
head(taxa.print)
```

#We now construct a phyloseq object directly from the dada2 outputs.
```{r}
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               tax_table(taxa))
ps
```

#Transform data to plot relative abundance.
```{r}
relative<- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
relative
```


#Graph relative abundance by phylum
```{r}
Phylum_graph <- plot_bar(relative, fill="Phylum", title="Relative Abundance of Phyla in Pumice Samples") +ylab("Relative Abundance")
Phylum_graph
```

#Print graph to a pdf file
```{r}
pdf("Phylum_graph.pdf")
Phylum_graph
dev.off()
```

#Graph relative abundance by Order
```{r}
Order_graph <-plot_bar(relative, fill="Order", title="Relative Abundance of Orders in Pumice Samples") +ylab("Relative Abundance") + theme(legend.position = "right") + guides(fill=guide_legend(ncol=2))
Order_graph
```
#Export graph as pdf
```{r}
pdf("Order_graph.pdf", width=40, height=50)
Order_graph
dev.off()
```

#Another option to export graphs
```{r}
ggsave("order_graph_gg.jpeg", plot=Order_graph, width=10, height=20, limitsize = FALSE )
```


#attempting to find shared ASVs
```{r}
#filter_taxa(ps, function(x) sum(x >= 1) == (2), TRUE)
#commonASVs = filter_taxa(ps, , TRUE)
#otu_table(commonASVs)

#otu_table(ps)[1:4, 1:5]
#merged<-merge_samples(ps, sample_names)
#filter<-filter_taxa(ps, function(x) sum(x > 1) >= (0*length(x)), TRUE)
#filter
#otu_table(filter)[1:4, 1:5]
```
#```{r}
#write.table(tax_table(ps), file="tax_output.csv")
#```
